\documentclass{article}
\pagestyle{headings}

% some of the packages used are
%\usepackage[utf8]{inputenc}
\usepackage{latexsym}
\usepackage{eepic}
\usepackage{makeidx}
\usepackage[dvips]{graphicx}
\usepackage[utf8]{inputenc}  %för ä ö å ?
\usepackage[english]{babel}
%\usepackage{times}
\usepackage{amssymb}
\usepackage{fancybox} 
\usepackage{textcomp}
\usepackage{float}
\usepackage{xcolor,colortbl}
\usepackage{tabularx}
\usepackage{enumitem}

\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
\definecolor{Gray}{gray}{0.85}
\definecolor{Orange}{rgb}{1,0.88,0}
\definecolor{LightCyan}{rgb}{0.7,0.9,0.9}
\definecolor{bluegray}{rgb}{0.4, 0.6, 0.8}
\definecolor{buff}{rgb}{0.94, 0.86, 0.51}
\definecolor{corn}{rgb}{0.98, 0.93, 0.36}
\definecolor{coralred}{rgb}{1.0, 0.25, 0.25}
\definecolor{coral}{rgb}{1.0, 0.5, 0.31}
\definecolor{applegreen}{rgb}{0.55, 0.71, 0.0}
\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}
\newcolumntype{g}{>{\columncolor{corn}}c}
\newcommand{\colcell}{\cellcolor{bluegray}}
\newcommand{\colcelltwo}{\cellcolor{coralred}}
\newcommand{\colcellthree}{\cellcolor{applegreen}}

% the fancy header/footer
% consult http://research.cmis.csiro.au/gjw/tex/docs/fancyhdr.pdf
% or some other fancy header documenatation for more info
\usepackage{fancyhdr}
\pagestyle{fancy}




\usepackage{blindtext}
\usepackage[utf8]{inputenc}
 
\title{Implementation Plan for LNS heuristik}
\author{Claes Arvidson, Emelie Karlsson}
\date{\today}
\begin{document}
 
\maketitle
 
\section*{First version \index{IMPL1}}
	\subsection*{Requirements}
	\begin{itemize}
	\item Implement a 5 week schedule
	\item Library on Wheels not considered
	\item Softer values such as PL, free day or HB preferences not considered
	\item Meetings not considered
	\item Objective function:
		\begin{itemize}
		\item Minimize unfulfilled worker demand (relaxed hard constraint)
		\item Maximize number of stand in librarians
		\item Maximize number of stand in assistants
		\end{itemize}
	\item Hard constraints:
		\begin{itemize}
		\item 1 task/day for a worker
		\item worker must be available to perform task
		\item worker must be qualified for task (lib/ass)
		\item the total number of tasks to be performed is constant
		\end{itemize} 
	\end{itemize}
	\subsection*{Implementation: task based approach} 
	\begin{itemize}
	\item Initial solution. First distribute weekends to obtain a feasible weekend schedule. Secondly distribute tasks according to some rule (for example by finding  the critical task hours). Initial solution not necessarily feasible with respect to the worker demand at each shift.
	\item Possible classes:
		\begin{itemize}
		\item Library (contains the demand schedule obtained from the data file, objective function costs associated with demand)
		\item Worker (contains availability (week blocks), identity, preferences etc., resulting task schedule, objective function costs associated with the worker)
		\end{itemize}
	\end{itemize}
	\subsection*{Implementation: week block based approach}
	\begin{itemize}
	\item Initial solution. Generate a schedule by assigning a feasible solution for each worker. The resulting schedule will most likely violate the demand constraint that is relaxed into the objective function.
	
	\item Classes in consideration:
		\begin{itemize}
		\item $Library$ (Information about current demand, total demand, workers assigned at a day $d$ and objective functions such as number of stand-ins are stored here)
		\item $Worker$ (contains all personal information about demands, qualifications, assigned blocks, blocks available for assignment etc.)
		\item $Block$ (a week long block containing an ID, type of block (weekend, weekrest, normal), number of tasks, tasks with information of day, shift and tasktype etc)
		\item $Pair\_of\_blocks$ (used later when implementing for 10 week period, containts information of $shift\_differ$ and which two blocks that forms the pair)
		\end{itemize}
		
	\item Destroy:
		\begin{itemize}
		\item Destroy week 1 for all workers (10 \% destroy) - need more than one week to move weekends (and weekrest).
		\item Destroy five consecutive weeks for around a fifth of the workers
		\item Identify and order the most problematic weeks for all workers and destroy in the order. Add some randomization as well to prevent getting stuck on a plateau.
		\end{itemize}
	\item Repair:
		\begin{itemize}
		\item Generate a new schedule with regard to number of tasks destroyed. The repair function searches through the pool of blocks available for each person and creates a new solution using a heuristic.
		\item The heuristic takes the demand and availability for a person into account. If a worker is available for a whole day - minimize the number of tasks given that day in order to generate a stand-in. 
		\end{itemize}
	\end{itemize}
	\section*{Second version \index{IMPL2}}
	\begin{itemize}
		\item Implement a 10 week schedule
		\item Library on Wheels considered
		\item Meetings considered
		\item Objective function:
				\begin{itemize}
				\item Minimize unfulfilled worker demand (relaxed hard constraint)
				\item Maximize number of stand in librarians
				\item Maximize number of stand in assistants
				\item Maximize similar weeks for workers
				\end{itemize}
				
	\end{itemize}
	
	\end{document}

